---
title: "Module 03 - Homework"
author: "Ethan Gutknecht"
date: "2/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



---------------------------



# Question 1

In a study of contamination of farm wells, contaminated water was classified as either “low contamination” or “high contamination,” and the distance of the well from a potential source of organic contamination was classified as either “nearby” or “not nearby.” Test for a significant association between distance and contamination using Fisher’s Exact Test for 2 X 2 Contingency Table. Use  = 0.05.

### Part I:
**Why might Fisher’s exact test for 2 X 2 contingency table be preferred to the two independent population proportions z-test or the chi-square test of homogeneity?**

For the study of farming wells, we can see that the expected counts of two of the four cells are below the minimum of five observations. Though, this is not a requirement, there is a requirement to have 15 successes and 15 failures. Since we do not meet this requirement, we should perform the Fisher's exact test.


### Part II:
**Label the parameters.**

Let p_drug = the true probability of there being an association between distance and contamination.

Let p_standard = the true probability of there being an association between distance and contamination.



### Part III:
**State null and alternative hypotheses.**

$H_{0}$: the distance does not affect the contamination.

$H_{a}$: the distance lowers the contamination.



### Part IV:
**Compute a test statistic and p-value using R.**
```{r}
# Create the table
table <- matrix(c(4, 3, 9, 0), nrow = 2)

# Perform fisher test
fisher.test(table, alternative="less")
```

### Part V:
**State the conclusion of the hypothesis test in the context of the problem.**

Based on the results of our test, we fail to reject the null hypothesis. This is because our p-value of 0.0625 is above our alpha value of 0.05. Thus, there is no correlation between distance and the contamination measurement of farming wells.





