---
title: "Module07InClass"
author: "Ethan Gutknecht"
date: "4/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1:
NarStor is a computer disk drive manufacturer. Technicians ran 16 NarStor hard drives continuously for almost three years. The times to failure (in hours) are given  in  the  following  table.  Construct  and  interpret  a  90%  bootstrap  (percentile)  confidence  interval for the mean failure time (in hours) of NarStor computer disk drive.
```{r}
data <- c(330,  620,  1870,  2410,  4620,  6396,  7822,  8102,  8309,  12882,  14419,  16092,  18384,  20916, 23812, 25814)
```

#### Part I:
Why  might  the  bootstrap  confidence  interval  be  preferred  to  the  one-sample  t-confidence interval for these data? 
```{r}
library(ggpubr)

par(mfrow=c(1,2)) # sets up a plot window for two side-by-side plots
qqnorm(data)
qqline(data)


hist(data, main = "Histogram of Times To Failure in Hours", xlab = "Time in Hours", ylab = "Frequency")

ggdensity(data, xlab = "Time in Hours", ylab= "Relative Frequency", main = "Density Plot of Hours")
```
The bootstrap method will be perfered 


#### Part II:
Label the parameter 
Let Mu be the mean failure time in hours of NarStor's computer disk drive.


#### Part III:
Construct a 90% bootstrap (percentile) confidence interval for the mean failure time (in hours) of NarStor computer disk drive using R
```{r}
bootmean <- numeric(1000)
for (i in 1:1000) {
    bootsample <- sample(data, size=length(data), replace=TRUE) 
    bootmean[i] <- mean(bootsample)
}
quantile(bootmean, c(0.05, 0.95))
```

#### Part IV:
Interpret  a  90%  bootstrap  (percentile)  confidence  interval  for  the  mean  failure time (in hours) of NarStor computer disk drive in context 
```{r}
par(mfrow=c(1,2))
hist(bootmean)
abline(v=mean(bootmean))
qqnorm(bootmean)
qqline(bootmean)
```

#### Part V:
Comment the overall shape of the sampling distribution of ð‘¥ (as approximated by bootstrapping) 

The overall shape of 


## Question 2
The following high temperature data were collected for randomly sampled days  in  the  summer  in  Des  Moines  and  Spokane.  Use  bootstrapping  to  estimate,  with  95% confidence, the true mean difference in high temperature in summer between Des Moines and Spokane. 
```{r}
A <- c(83, 91, 94, 89, 89, 96, 91, 92, 90)
B <- c(78, 82, 81, 77, 79, 81, 80, 81)
```

#### Part I:
Why might the bootstrap confidence interval be preferred to the two independent random samples t-confidence interval for population means for these data?
```{r}
par(mfrow=c(1,2))

hist(A, main = "Histogram of Diet A", xlab = "Growth", ylab = "Frequency")
qqnorm(A)
qqline(A)
ggdensity(A, xlab = "Growth", ylab= "Relative Frequency", main = "Density Plot of Diet A")

hist(B, main = "Histogram of Diet B", xlab = "Growth", ylab = "Frequency")
qqnorm(B)
qqline(B)

ggdensity(B, xlab = "Growth", ylab= "Relative Frequency", main = "Density Plot of Diet B")
```

#### Part II:
Label the parameters.

Let Ma be the mean high temperature data of summer days in Des Moines
Let Mb be the mean high temperature data of summer days in Spokane.

Let (Ma - Mb) be the difference between the two cities.

#### Part III:
From  the  bootstrap  distribution  for  the  sample  mean  differences,  calculate  the estimated bias. Comment the overall shape of the sampling distribution of sample mean  differences (as approximated by bootstrapping) 
```{r}
obs.diff <- mean(A) - mean(B)
bootdiff <- numeric(1000)
for (i in 1:1000) {
  bootsampleA <- sample(A, size=length(A), replace=TRUE)
  bootsampleB <- sample(B, size=length(B), replace=TRUE)
  bootdiff[i] <- mean(bootsampleA) - mean(bootsampleB)
}

est.bias <- mean(bootdiff) - obs.diff
est.bias

par(mfrow=c(1,2))
hist(bootdiff)
abline(v=obs.diff)
abline(v=mean(bootdiff),lty=2) # add dashed line mean of the bootstrap dist
qqnorm(bootdiff)
qqline(bootdiff)
```

#### Part IV:
Construct and interpret a 95% bootstrap confidence interval (BCa based) for the  true mean difference in high temperature in summer between Des Moines and Spokane using R add-on package boot 

```{r}
library(boot)

mymean <- function(bootdiff,i) mean(bootdiff[i])
myboot <- boot(bootdiff, mymean, R = 1000)
myboot
```




