---
title: "Module08Notes"
author: "Ethan Gutknecht"
date: "4/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Module 08 - Notes
## Review of Simple Linear Regression
Question 1:
A person’s muscle mass is expected to decrease with age. To explore this relationship in women, a nutritionist randomly selected 15 women from each 10-year age group beginning with age 40 and ending with age 79. The data reside in an excel sheet Muscle. The variables in the dataset are Age and Mass.

###### Part I
Investigate the relationship between age and muscle mass visually using a scatterplot. Describe scatterplot

```{r eval=FALSE}
Muscle <- read.csv("Muscle.csv", header = TRUE)

plot(Mass~Age, data = Muscle, main ="Muscle Mass vs Age")
```

###### Part II:
Using R, estimate the regression parameters. Report the regression equation using variable names. Interpret the slope of the regression line in context.

```{r eval=FALSE}
fit <- lm(Mass~Age, data = Muscle)
summary(fit)
```

###### Part III:
Find and interpret the 95% confidence interval for mean muscle mass in 65 years old women

```{r eval=FALSE}
predict(fit, newdata = data.frame(Age = 65), int ="conf")
```

###### Part IV:
Overlay the regression line on the scatterplot
```{r eval=FALSE}
abline(fit, col = "Red")
```




---------------------------------------
---------------------------------------
---------------------------------------




## Assessing the SLR Model Fit

#### Assumptions:
- Linearity (L)
- Independence (I) - Collected randomly
- Normality (N) - QQPlot, histogram of residuals
- Equality of variance (E)


There is a negative linear strong association between the age of muscle mass in women aged 40-79. As age increases the muscle mass decreases. Ther is no distinct outlier present so a simple linear regression modle seems viable here.

0 <= r <= 0.4     WEAK
0.4 <= r <= 0.8   MODERATE
GREATER than 0.8  STRONG

###### Part V:
The normal QQ plot of the residual suggests that any violation of normality assumption because of the points follow a diagonal straight line. However the residual plot shows some megaphone or fanning effect suggesting the equal variance assumption.


## Locally Weighted Scatterplot Smoothing
Question 2: 
A high respiratory rate is a potential diagnostic indicator of respiratory infection in children. To judge whether a respiratory rate is truly “high,” however, a physician must have a clear picture of the distribution of normal respiratory rates. To this end, Italian researchers measured the respiratory rates of n = 618 children between the ages of 15 days and 3 years (given in months). The data reside in an excel sheet Respiratory. The variables in the dataset are Age and Rate.

###### Part I:
Create residual plot of (X,e) and normal QQ plot of residual. Comment any assumption violations of simple linear regression

```{r eval=FALSE}
Respiratory <- read.csv("Respiratory", header = TRUE)


fit <- lm(Rate~Age, data = Respiratory)
par(mfrow=c(1,2))
X <- Respiratory$Age
e <- resid(fit)
plot(X,e,ylab = "Residuals", xlab="Age", main="Residual Plot of (X,e)")
abline(0,0, col = "Red")
qqnorm(e)
qqline(e)
```
Normal qq plot: a few points on the normal qq plot do nto follow the diagonal straight line this suggest a moderate voilation of normlaity assumoption.




###### Part II:
Fit a nonparametric LOESS regression curve to the data, and use it to calculate a 95% bootstrap confidence interval for the true mean respiratory rate for children aged 12 months. Interpret the bootstrap confidence interval for the true mean respiratory rate for children aged 12 months.

```{r eval=FALSE}
par(mfrow=c(1,2)) # set up two panels for our plots
bootpred <- numeric(1000) # create vector to hold bootstrapped predictions
hyp.age <- 12 # set the hypothesized x value for prediction
rate.lo <- loess(Rate~Age, data = Respiratory)
pred <- predict(rate.lo, data.frame(Age = hyp.age))
```

We can be 95% confident that the true mean respitary rate for children



###### Part III:
Generate a picture of the bootstrapped LOESS curves
```{r eval=FALSE}
x <- Respiratory$Age
y <- Respiratory$Rate
out <- loess(y~x, span = 0.3)
plot(y~x, main="1000 bootstrapped LOESS curve fits")
curve(predict(out, newdata=data.frame(x = x)), add=TRUE)
n <- length(x)

nboot <- 1000
for (i in 1:nboot) {
index <- sample(n, replace = TRUE)
boot.x <- x[index]
boot.y <- y[index]
boot.out <- loess(boot.y~boot.x, span=0.3)
curve(predict(boot.out, data.frame(boot.x = x)), add=TRUE,
col="red")
}
points(x, y)
curve(predict(out, newdata=data.frame(x = x)), add=TRUE, lwd=2)
```


###### Part IV:
Find the same prediction (i.e. a 95% CI for the true mean respiratory rate in 12 months old children), only now using a conventional simple linear regression. How does the result compare to your bootstrap CI found from a LOESS model?
```{r eval=FALSE}
slrfit <- lm(Rate~Age, data = Respiratory)
predict(slrfit, newdata=data.frame(Age=12), int="conf")
```

true mean respitory rate for 12 month olds of (38.08, 39.33). The loess model with bootstraping yeilds (35.17, 37.49).

SLR(38.08, 39.33)
Loess (35.17, 37.49)









