---
title: "Module02 - In Class"
author: "Ethan Gutknecht"
date: "2/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggpubr)
library(exactRankTests)
library(randtests)
```

# Module 02 - In Class Exercise


## Question 1
#### Part 1

```{r}
q1Data <- c(7.1, 4.4, 3.9, 2.2, 4.2, 3.4, 1.1, 6.1, 2.2, 6.7, 4.9, 7.3, 0.3, 7.6)

hist(q1Data,  main  =  "Histogram  of  Distances",  xlab  =  "Distance  in 
Miles", ylab = "Frequency")

ggdensity(q1Data,  xlab  ="Distance To Emergency", 
ylab="Relative Frequency", main="Density Plot of Hour Spend")
```

Since this data does not seem to be normal, it breaks the assumption. Due to this, we will have to find the median using a nonparametric statistic. The best test that we have that fits this questions needs is the Wilcoxon Signed-Rank test.



#### Part II 
m = median accident distance from the dam



#### Part III
$H_{0}$: m = 4

$H_{a}$: m < 4



#### Part IV
```{r warning=FALSE}
wilcox.exact(q1Data, alternative = c("greater"), mu = 4);
```



#### Part V
Since the P-value is greater than 0.05, we have insufficient evidence from the test that states taht the median is greater than four.



## Question 2
#### Part I

$H_{0}$: There is no trend in the number of crops sold.

$H_{a}$: There is a downward trend in the number of crops sold.



#### Part II

```{r}

# Create dataframe
q2Data <- data.frame(
  CropsInsured = c(19479, 26667, 63969, 57715, 38086, 38434, 24196, 19319, 29975, 25451, 20410, 19910, 15628, 15375, 21312, 26526, 24865, 21152, 23458, 25774, 32646, 31786, 24821, 19593, 14960),
  Years = c(1948, 1949, 1950, 1951, 1952, 1953, 1954, 1955, 1956, 1957, 1958, 1959, 1960, 1961, 1962, 1963, 1964, 1965, 1966, 1967, 1968, 1969, 1970, 1971, 1972))

# Compute test
cox.stuart.test(q2Data$CropsInsured, "left.sided") 
```


#### Part III
The test fails to reject the null hypothesis due to the p-value being under 0.05.
